```{r reprex-options, include = FALSE}
options(
  keep.source = TRUE,
  rlang_backtrace_on_error_report = "full",
  crayon.enabled = FALSE,
  reprex.current_venue = "gh"
)
```

```{r, results = 'asis', echo = FALSE, include = file.exists('.Rprofile'), eval = file.exists('.Rprofile')}
cat(sprintf("*Local `.Rprofile` detected at `%s`*", normalizePath(".Rprofile")))
```

---
output:
  reprex::reprex_document:
    venue: "gh"
    advertise: FALSE
    session_info: TRUE
    style: TRUE
    comment: "#;-)"
    tidyverse_quiet: FALSE
    std_out_err: TRUE
knit: reprex::reprex_render
---

This template demonstrates many of the bells and whistles of the `reprex::reprex_document()` output format. The YAML sets many options to non-default values, such as using `#;-)` as the comment in front of output.

## Code style

Since `style` is `TRUE`, this difficult-to-read code (look at the `.Rmd` source file) will be restyled according to the Tidyverse style guide when it's rendered. Whitespace rationing is not in effect!

```{r}
x <- 1
```


<details style="margin-bottom:10px;">
<summary>Standard output and standard error</summary>
`C:/Users/Eriba/OneDrive/UMCG/E-Cadherin Project/IDIRT/Untitled-2_reprex_std_out_err.txt`
</details>

<details style="margin-bottom:10px;">
<summary>Session info</summary>
```{r }
sessioninfo::session_info()
```
</details>
