---
title: "I-DIRT Data Analysis: iDIRT set"
output: html_document
---

## Set-up

```{r}
# Object preparation From Proteomics Only
iDIRT_set <- list(
  "Path_iDIRT_file" = "../data/IDIRT data/IDIRTall_Proteins.txt",
  "Metadata_iDIRT_file" = "../data/Metadata/iDIRT_metadata.csv" 
)

# Libraries
library(tidyverse)
```


## Data Preparation

1. Load iDIRT data
2. Apply Filters

  * Protein.FDR.Confidence.Combined == High
  * Master == IsMasterProtein
  * Number.of.Unique.Peptides > 1

```{r}
## From Proteome Discover
idirt_data <- read.delim(iDIRT_set[["Path_iDIRT_file"]],
                         header = TRUE, stringsAsFactors = FALSE) %>% 
  ## Cleaning data
  filter(
    Protein.FDR.Confidence.Combined == "High",
    Master == "IsMasterProtein",
    Number.of.Unique.Peptides > 1
  )

dim(idirt_data)
```

3) From Metadata (IDIRT Metadata) select the columns of interest. Working with NO SWAP data. Working with Buffer 02

4) Filter to remove the rows with NA except in at least one replicates (including channel)

5) Calculate ratio

$$ Ratio = Sample_x / (Sample_x + Control_x) $$

6) Calculate median of Ratios (replicates)

7) Calculate Median Abundance Stoichiometry

   Get the median Sample by protein
   Get the median Control by protein
   Get the median in Control
   Select the protein of interest
   Calculate Median in Sample (POI) / Number of Peptides 
   Median Sample (Proteins)/Number of Peptides/ POI
   

```{r}
# Load metadata
idirt_metadata <- read.csv(iDIRT_set[["Metadata_iDIRT_file"]], 
                           sep = ",", stringsAsFactors = FALSE)

## Select NO SWAP and BUFFER 02
Exp_sel <-  idirt_metadata %>% 
  filter(SWAP == "NO") 

## Separate by Buffer the iDIRT data
idirt_Stats <- list()
idirt_ratios <- list()
for (i in unique(Exp_sel$Buffer)) {
  #Select columns of interest
  col_exp <- Exp_sel %>% 
    filter(Buffer == i) %>% 
    pull(Col_Name)
  
  # Select Proteins Metadata
  idirt_prot_meta <- idirt_data %>% 
    dplyr::select(Accession, Number.of.Peptides)
  
  # Select from IDIRT
  df <- idirt_data %>% 
    dplyr::select(Accession, all_of(col_exp))

  ##--->Still working with each table
  # Remove proteins with NA
  # Replace Missing values with zero
  temp <- df %>% 
    filter(rowSums(is.na(df[,-1])) < ncol(df) -1) %>% 
    replace(is.na(.), 0)
  
  ## Split Control and Sample to calculate ratio
  ## CONTROL Group
  control_col <- Exp_sel %>% 
    filter(Buffer == i, Condition == "Control") %>% 
    pull(Col_Name)
  
  control_df <- temp %>% dplyr::select(all_of(control_col))
  
  ## SAMPLE Group
  sample_col <- Exp_sel %>% 
    filter(Buffer == i, Condition != "Control") %>% 
    pull(Col_Name)
  
  sample_df <- temp %>% dplyr::select(all_of(sample_col))
  
  ## Calculate ratio as a matrix
  ratios_df <- sample_df/(control_df+sample_df)
  
  colnames(ratios_df) <- paste0("iDIRT_rep", c(1:ncol(ratios_df)))

  ratios_df$Accession <- temp$Accession
  
  # Calculate median Ratios
  idirt_summary <- ratios_df %>% 
    pivot_longer(cols =-Accession, names_to = "iDirt", values_to = "Ratio") %>% 
      group_by(Accession) %>% 
      summarize(
        IDIRT_mean = mean(Ratio, na.rm = TRUE),
        IDIRT_sd = sd(Ratio, na.rm = TRUE),
        IDIRT_median = median(Ratio, na.rm = TRUE),
        IDIRT_Missing = sum(is.na(Ratio))
            ) %>% 
    mutate(Experiment = paste0("Buffer_",i))

  
  ## Summarize Abundances in Sample and Control
  
  Channels_summary <- temp %>% 
    pivot_longer(cols = -Accession,
                 names_to = "Experiment", values_to = "Values") %>% 
    mutate(Channel = ifelse(grepl("light", Experiment, ignore.case = TRUE),
                            "Light", "Heavy")) %>% 
    group_by(Accession, Channel) %>% 
    summarize(
        Channel_mean = mean(Values, na.rm = TRUE),
        Chanel_sd = sd(Values, na.rm = TRUE),
        Channel_median = median(Values, na.rm = TRUE),
        Channel_Missing = sum(is.na(Values))
    ) %>% 
    pivot_wider(id_cols = Accession, names_from = Channel, values_from = -c(Accession, Channel)) %>% 
    mutate(Experiment = paste0("Buffer_", i)) %>% 
    left_join(idirt_prot_meta, by = "Accession")
  
  ## Calculate Median Abundance Stoichiometry
  #control_median <- median(Channels_summary$Channel_median_)
  
  # Protein of Interest
  POI_Stoichiometry <- Channels_summary %>% 
    filter(Accession == "P12830") %>% 
    mutate(Stoichiometry = Channel_median_Heavy/Number.of.Peptides) %>% 
    pull(Stoichiometry)
  
  Channels_summary$abundance_stoichiometry <-  Channels_summary$Channel_median_Heavy/Channels_summary$Number.of.Peptides/POI_Stoichiometry
  
  idirt_Stats[[paste0("Buffer_", i)]] <- Channels_summary %>% ungroup()
  
  
  idirt_ratios[[paste0("Ratios_Buffer_", i)]] <- ratios_df %>% 
    left_join(idirt_summary, by = "Accession") %>%
    ungroup()
  
}
```


## Data Quality

Compare Light and Heavy abundance change:

* Correlation and Distribution

```{r}
# Gather all the data
Idirt_stats_all <- do.call(rbind, idirt_Stats)
head(Idirt_stats_all)
Idirt_stats_all %>% 
  filter(Accession == "A0A024QZ42")
```

```{r}
# Scattter plot between experiments
ggplot(Idirt_stats_all, aes(x = log2(Channel_median_Heavy +1),
                            y = log2(Channel_median_Light+1),
                            color = Experiment,
                            size = abundance_stoichiometry)) +
  geom_point(show.legend = FALSE, alpha = 0.6) +
  facet_wrap(.~Experiment) +
  theme_bw()
```


## Ratio Missing 

```{r}
# Gather all the ratio data
IdirtRatios_stats_all <- do.call(rbind, idirt_ratios)
head(IdirtRatios_stats_all)
```


```{r}
counts_missing <- IdirtRatios_stats_all %>% 
  count(Experiment, IDIRT_Missing)

ggplot(counts_missing, aes(x = IDIRT_Missing, y = n, fill = Experiment)) +
  geom_bar(show.legend = FALSE, stat = "identity") +
  geom_text(aes(x = IDIRT_Missing, y = n + 20, label = n), size= 2) +
  facet_wrap(.~Experiment) +
  theme_bw()

```

7) Keep proteins that contain values in at least 2 replicates 

```{r}
IdirtRatios_filter <- IdirtRatios_stats_all %>% 
  filter(IDIRT_Missing < 3)
```

## Get Stoichiometry


## Comparison with Proteomics Mix

* Shared proteins

```{r}
idirt_proteins_byBuffer <- lapply(idirt_ratios, function(x) unique(x$Accession))

# Use upset
library(UpSetR)

comparisons_list <- lapply(names(idirt_proteins_byBuffer), function(x) {list("Proteomics", x)})
## Add the Proteomics
idirt_proteins_byBuffer$Proteomics <- unique(Proteomics_Ratio_All$Accession)

upset(fromList(idirt_proteins_byBuffer), nsets = 7)
```

In euler

```{r}
library(eulerr)

Euler_plots <- list()
for (i in grep("Ratios",names(idirt_proteins_byBuffer), value = TRUE)) {
  name_sample <- i
  x <- list(
    a =idirt_proteins_byBuffer[[i]],
    b = idirt_proteins_byBuffer$Proteomics
          )
  
  names(x) <- c(i, "Proteomics")

  p <- plot(euler(x, shaple = "ellipse"), labels = list(cex=0.5),
            quantities = list(cex = 0.5))
  
  Euler_plots[[i]] <- p

}
```


```{r}
Euler_plots$Ratios_Buffer_2
```

**Check Ratio in Mix**

```{r}
# Get the Ratio median for each protein
Proteomics_summary <- Proteomics_Ratio_All %>% 
  group_by(Accession) %>% 
  summarise(
    Ratio_mean = mean(Ratio, na.rm = TRUE),
    Ratio_median = median(Ratio, na.rm = TRUE),
    Ratio_sd = sd(Ratio, na.rm = TRUE),
    Ratio_missing = sum(is.na(Ratio)),
    # Control
    Control_mean = mean(Control, na.rm = TRUE),
    Control_median = median(Control, na.rm = TRUE),
    Control_sd = sd(Control, na.rm = TRUE),
    Control_missing = sum(is.na(Control)),
    # Sample
    Sample_mean = mean(Sample, na.rm = TRUE),
    Sample_median = median(Sample, na.rm = TRUE),
    Sample_sd = sd(Sample, na.rm = TRUE),
    Sample_missing = sum(is.na(Sample)),
  )
 

## Add the proteomics Ratio to the iDIRT
IDIRT_Mix <- IdirtRatios_filter %>% 
  left_join(Proteomics_summary, by = "Accession")
head(IDIRT_Mix)
```

Replace Missing values From `Mix` Data with -0.15. **Why this value?**.

Add Classification:
  * if IDIRT_ratio_median

```{r}
IDIRT_Mix <-IDIRT_Mix %>%
  mutate(
    Ratio_median = if_else(is.na(Ratio_median), -0.15, Ratio_median),
    # Add groups
    Group_iDIRT = if_else(IDIRT_median>0.71, "High", "Low")
    )
head(IDIRT_Mix)
```

## Quality Parallel
 
```{r}
# library(GGally)
# 
# ggparcoord(IDIRT_Mix, 
#            col = c(8, 12),
#            scale = "globalminmax",
#            groupColumn = 24)
```
 
```{r}
# Without GGally
temp <- IDIRT_Mix %>% 
  dplyr::select(Accession, Group_iDIRT, IDIRT_median, Ratio_median, Experiment) %>% 
  pivot_longer(cols = -c(Accession, Group_iDIRT, Experiment), 
               names_to = "Proteo_Interact",
               values_to = "Ratio")

head(temp)
ggplot(temp) +
  geom_line(aes(x = Proteo_Interact, y = Ratio, color = Group_iDIRT, group = Accession),
            alpha = 0.3)+
  geom_point(aes(x = Proteo_Interact, y = Ratio, color = Group_iDIRT),
             alpha = 0.25) +
  geom_hline(yintercept = 0.71, color = "blue", linetype = "dashed") +
  facet_wrap(.~Experiment) +
  theme_classic() +
  theme(
    legend.position = "none"
  )
```


Ratio Distribution

```{r}
ggplot(temp) +
  #geom_histogram(aes(x = Ratio, fill = Proteo_Interact),
  #            alpha = 0.4, binwidth = 0.025) +
  geom_density(aes(x = Ratio, fill = Proteo_Interact),
                alpha = 0.4) +
  facet_wrap(.~Experiment) +
  theme_classic()
```

Alternative could be box plot


```{r}
ggplot(temp) +
  geom_violin(aes(x = Ratio, y = Proteo_Interact, color = Proteo_Interact),
               alpha = 0.4) +
  geom_point(aes(x = Ratio, y = Proteo_Interact, color = Proteo_Interact),
             size = 0.75,
             alpha = 0.4, 
             position = position_jitter(height = 0.2)) +
  facet_wrap(.~Experiment) +
  theme_classic()
```

Plot abundance and Ratio

Use the data that contains the Abundance and Ratios. 
From the Light Channel data, remove all the proteins that contains Zero.

Get the quantile by Experiment

```{r}
temp_stats <- IdirtRatios_filter %>% 
  left_join(Idirt_stats_all, by = c("Accession", "Experiment"))
head(temp_stats)
```


```{r}
# Remove the protiens with zero in Heavy?
# Why not remove proteins with zero in Both Channels
temp_stats %>% 
  filter(Channel_median_Heavy > 0) %>% 
  ggplot(aes(x = reorder(Accession, -Channel_median_Heavy), 
             y = log2(Channel_median_Heavy+ 1))) +
  geom_point(color = "69b3a2", alpha = 0.5) +
  ## Plot the IDIRT significant
  geom_point(data = temp_stats %>% filter(IDIRT_median >= 0.71),
             aes(x = Accession, 
                 y = log2(Channel_median_Heavy+ 1)),
             color = "red") +
  facet_wrap(.~Experiment)
```

The plot is rank ratio against the abundance in Proteomics

```{r}
temp_rank <- temp_stats %>% 
  filter(Channel_median_Light > 0,
         Channel_median_Heavy > 0) %>% 
  group_by(Experiment) %>%
  mutate(rank = rank(Channel_median_Heavy, ties.method = "first"))

temp_quantiles <- temp_stats %>% 
  group_by(Experiment) %>% 
  summarize(Quantile_2 = log2(quantile(Channel_median_Light)[2]))

temp_rank %>% 
  ggplot(aes(x = rank, 
             y = log2(Channel_median_Heavy+ 1))) +
  geom_point(color = "#69b3a2", alpha = 0.5) +
  ## Plot the IDIRT significant
  geom_point(data = temp_rank %>% filter(IDIRT_median >= 0.71),
            aes(x = rank, 
                y = log2(Channel_median_Heavy+ 1)),
             color = "darkred") +
  # Show the protein of interest
  ggrepel::geom_text_repel(data = temp_rank %>% filter(Accession == "P12830"),
                  aes(label = Accession), size = 2.5)+
  # Plot quantiles
  geom_hline(data = temp_quantiles, 
             aes(yintercept = Quantile_2),
             linetype = "dashed", color = "red") +
  labs(x = "Rank", y = "log2(Abundance heavy labeled proteins + 1)") +
  facet_wrap(.~Experiment) +
  theme_classic() +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )
```


Return the data frame with the IDIRT Ratios and after filtering (proteins in at least 2 replicates).

```{r}
saveRDS(IdirtRatios_filter, file = "IDIRT_Loaded.Rds")
```

